---
title: "EDA"
author: "mohsin"
date: "27 January 2017"
output:
  html_document: default
  pdf_document: default
---

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
```


```{r}
train_data <- read.csv("train.csv")
test_data <- read.csv("test.csv")
str(train_data)

calc_corr <- function(var_data){
  indices <- sapply(var_data, function(x) !is.na(x))
  clean_data <- var_data[indices]
  prices <- train_data$SalePrice[indices]
  cor(clean_data, prices)
}

hist_category <- function(var_data) {
  ggplot(data.frame(var_data), aes(x=var_data)) + geom_bar()
}

summary_category <- function(var_data){
  cat(paste0("levels : ", levels(var_data), "\n"))
  hist_category(var_data)
  ggplot(train_data, aes(x=var_data, y=SalePrice)) + geom_boxplot()
}

plot_price_vs_var <- function(var_data){
  ggplot(train_data, aes(x = var_data, y=SalePrice)) +geom_point()
}

summary_cont <- function(var_data){
  cat("summary is ")
  cat(summary(var_data))

  boxplot(var_data)

  # Correlation with the price
  cat("Correlation with the sale price is ",calc_corr(var_data))
}
```

## Univariate analysis

1) ID -lite
### 2) MSSubClass : The building class
```{r}
factor_mssubclass_levels <- as.factor(train_data$MSSubClass)

levels(factor_mssubclass_levels)
ggp <- ggplot(data.frame(factor_mssubclass_levels) , aes(x=factor_mssubclass_levels))
ggp + geom_bar()

ggplot(train_data, aes(x=factor_mssubclass_levels, y=SalePrice)) + geom_boxplot()
```

The data is not uniformly distributed

3) MSZoning : The general zoning classification

```{r}
levels(train_data$MSZoning)
ggplot(data.frame(train_data$MSZoning), aes(x=train_data$MSZoning)) + geom_bar()
ggplot(train_data, aes(x=MSZoning, y=SalePrice)) + geom_boxplot()
```

### 3) Lot Frontage
```{r}
summary(train_data$LotFrontage)

boxplot(train_data$LotFrontage)
ggplot(train_data, aes(x = LotFrontage, y=SalePrice)) +geom_point()

# Correlation with the price
calc_corr(train_data$LotFrontage)
```

### 4) Lot area : Linear feet of street connected to property

```{r, include=FALSE}
summary(train_data$LotArea)
ggplot(train_data, aes(x = LotArea, y=SalePrice)) +geom_point()

boxplot(train_data$LotArea)

# Correlation with the price
calc_corr(train_data$LotArea)
```

```{r}
plot_price_vs_var(train_data$LotArea)
summary_cont(train_data$LotArea)
```

### 5) Street : Type of road access

```{r}
hist_category(train_data$Street)
summary_category(train_data$Street)
```

I think that the above varibale can be removed as no entropy is there in the data

### 6) Alley: Type of alley access
```{r}
var_data <- train_data$Alley
hist_category(var_data)
summary_category(var_data)
```

Again not required as we have lot of NAs

### 7) LotShape: General shape of property
```{r}
var_data <- train_data$LotShape
hist_category(var_data)
summary_category(var_data)
```

### 8) LandContour: Flatness of the property
```{r}
var_data <- train_data$LandContour
hist_category(var_data)
summary_category(var_data)
```

### 9) Utilities: Type of utilities available
```{r}
var_data <- train_data$Utilities
hist_category(var_data)
summary_category(var_data)
```
This can also be removed as no entropy in the data

### 10) LotConfig: Lot configuration
```{r}
var_data <- train_data$LotConfig
hist_category(var_data)
summary_category(var_data)
```
This can also be removed as no entropy in the data

### 11) LandSlope: Slope of property
```{r}
var_data <- train_data$LandSlope
hist_category(var_data)
summary_category(var_data)
```
This can also be removed as no entropy in the data

### 12) Neighborhood: Physical locations within Ames city limits
```{r}
var_data <- train_data$Neighborhood
hist_category(var_data)
summary_category(var_data)
```

### 13) Condition1: Proximity to main road or railroad
```{r}
var_data <- train_data$Condition1
hist_category(var_data)
summary_category(var_data)
```
This can also be removed as no entropy in the data

### 14) Condition2: Proximity to main road or railroad (if a second is present)
```{r}
var_data <- train_data$Condition2
hist_category(var_data)
summary_category(var_data)
```
This can also be removed as no entropy in the data

### 13) BldgType: Type of dwelling
```{r}
var_data <- train_data$BldgType
hist_category(var_data)
summary_category(var_data)
```

### 14) HouseStyle: Style of dwelling
```{r}
var_data <- train_data$HouseStyle
hist_category(var_data)
summary_category(var_data)
```

Can be grouped int a binary variable with 1story as a variable and the rest as others

### 14) HouseStyle: Style of dwelling
```{r}
var_data <- train_data$HouseStyle
hist_category(var_data)
summary_category(var_data)
```


### 15) Overall quality: Overall material and finish quality

```{r}
plot_price_vs_var(train_data$OverallQual)
summary_cont(train_data$OverallQual)
```


### 16) Overall condition: Overall condition rating

```{r}
plot_price_vs_var(train_data$OverallCond)
summary_cont(train_data$OverallCond)
```

### 17) YearBuilt: Original construction date

```{r}
plot_price_vs_var(train_data$YearBuilt)
summary_cont(train_data$YearBuilt)
```

Need to scale by putting the smallest date as 0 .

### 18) YearRemodAdd: Remodel date

```{r}
plot_price_vs_var(train_data$YearRemodAdd)
summary_cont(train_data$YearRemodAdd)
```

Need to scale by putting the smallest date as 0 .

### 19) RoofStyle: Type of roof

```{r}
var_data <- train_data$RoofStyle
hist_category(var_data)
summary_category(var_data)
```

Can be binned nto gable and non gable

### 20) RoofMatl: Roof material

```{r}
var_data <- train_data$RoofMatl
hist_category(var_data)
summary_category(var_data)
```

No entropy in the data, so can be ignored

### 21) Exterior1st: Exterior covering on house

```{r}
var_data <- train_data$Exterior1st
hist_category(var_data)
summary_category(var_data)
```

### 22) Exterior2nd: Exterior covering on house (if more than one material

```{r}
var_data <- train_data$Exterior2nd
hist_category(var_data)
summary_category(var_data)
```

### 23) MasVnrType: Masonry veneer type

```{r}
var_data <- train_data$MasVnrType
hist_category(var_data)
summary_category(var_data)
```

Binary binning can be done. None and non none.

### 24) MasVnrArea: Masonry veneer area in square feet

```{r}
plot_price_vs_var(train_data$MasVnrArea)
summary_cont(train_data$MasVnrArea)
```

I think it will be advisable to use only one of the above two variables. Lets experiment.

### 25) ExterQual: Exterior material quality

```{r}
var_data <- train_data$ExterQual
hist_category(var_data)
summary_category(var_data)
```

Can be converted to numeric variables.

### 26) ExterCond: Present condition of the material on the exterior

```{r}
var_data <- train_data$ExterCond
hist_category(var_data)
summary_category(var_data)
```

Can be converted to numeric variables.

### 27) Foundation: Type of foundation

```{r}
var_data <- train_data$Foundation
hist_category(var_data)
summary_category(var_data)
```

### 28) BsmtQual: Evaluates the height of the basement

```{r}
var_data <- train_data$BsmtQual
hist_category(var_data)
summary_category(var_data)
```

Can be converted to numeric variables.

### 29) BsmtCond: Evaluates the general condition of the basement

```{r}
var_data <- train_data$BsmtCond
hist_category(var_data)
summary_category(var_data)
```

Can be converted to numeric variables. With Po as 0, Na as 1, Fa as 2, TA as 3 and Gd as 4

Bivariate analysis: Need to do cross tab

### 30) BsmtExposure: Refers to walkout or garden level walls

```{r}
var_data <- train_data$BsmtExposure
hist_category(var_data)
summary_category(var_data)
```

### 31) BsmtFinType1: Rating of basement finished area

```{r}
var_data <- train_data$BsmtFinType1
hist_category(var_data)
summary_category(var_data)
```

### 32) BsmtFinSF1: Type 1 finished square feet

```{r}
plot_price_vs_var(train_data$BsmtFinSF1)
summary_cont(train_data$BsmtFinSF1)
```

### 33) BsmtFinType2: Rating of basement finished area (if multiple types)

```{r}
var_data <- train_data$BsmtFinType2
hist_category(var_data)
summary_category(var_data)
```

### 34) BsmtFinSF2: Type 2 finished square feet

```{r}
plot_price_vs_var(train_data$BsmtFinSF2)
summary_cont(train_data$BsmtFinSF2)
```

### 35) BsmtUnfSF: Unfinished square feet of basement area

```{r}
plot_price_vs_var(train_data$BsmtUnfSF)
summary_cont(train_data$BsmtUnfSF)
```

### 36) TotalBsmtSF: Total square feet of basement area

```{r}
plot_price_vs_var(train_data$TotalBsmtSF)
summary_cont(train_data$TotalBsmtSF)
```

### 37) Heating: Type of heating

```{r}
var_data <- train_data$Heating
hist_category(var_data)
summary_category(var_data)
```

The entropy of the data is less and also I think heating quality is more important

### 38) HeatingQC: Heating quality and condition

```{r}
var_data <- train_data$HeatingQC
hist_category(var_data)
summary_category(var_data)
```

### 39) CentralAir: Central air conditioning

```{r}
var_data <- train_data$CentralAir
hist_category(var_data)
summary_category(var_data)
```

### 40) Electrical: Electrical system

```{r}
var_data <- train_data$Electrical
hist_category(var_data)
summary_category(var_data)
```

Might have to do some data imputation and binning can be done

### 41) 1stFlrSF: First Floor square feet
 
```{r}
plot_price_vs_var(train_data$X1stFlrSF)
summary_cont(train_data$X1stFlrSF)
```

### 42) 2ndFlrSF: Second floor square feet
 
```{r}
plot_price_vs_var(train_data$X2ndFlrSF)
summary_cont(train_data$X2ndFlrSF)
```

### 43) LowQualFinSF: Low quality finished square feet (all floors)
 
```{r}
plot_price_vs_var(train_data$LowQualFinSF)
summary_cont(train_data$LowQualFinSF)
```

### 44) GrLivArea: Above grade (ground) living area square feet
 
```{r}
plot_price_vs_var(train_data$GrLivArea)
summary_cont(train_data$GrLivArea)
```

### 45) BsmtFullBath: Basement full bathrooms

For EDA, I am using treating this variable as a category variable

```{r}
var_data <- train_data$BsmtFullBath %>% as.factor()
hist_category(var_data)
summary_category(var_data)
```

Does not look like very helpful variable but still worth a try

### 46) BsmtHalfBath: Basement half bathrooms

For EDA, I am using treating this variable as a category variable

```{r}
var_data <- train_data$BsmtHalfBath %>% as.factor()
hist_category(var_data)
summary_category(var_data)
```

Does not look like very helpful variable but still worth a try

### 47) FullBath: Full bathrooms above grade

For EDA, I am using treating this variable as a category variable

```{r}
var_data <- train_data$FullBath %>% as.factor()
hist_category(var_data)
summary_category(var_data)
calc_corr(train_data$FullBath)
```

Looks important variable

### 48) HalfBath: Half baths above grade

For EDA, I am using treating this variable as a category variable

```{r}
var_data <- train_data$HalfBath %>% as.factor()
hist_category(var_data)
summary_category(var_data)
calc_corr(train_data$HalfBath)
```

### 49) Bedroom: Bedrooms above grade

For EDA, I am using treating this variable as a category variable

```{r}
var_data <- train_data$BedroomAbvGr %>% as.factor()
hist_category(var_data)
summary_category(var_data)
calc_corr(train_data$BedroomAbvGr)
```

### 50) Kitchen: Kitchens above grade

For EDA, I am using treating this variable as a category variable

```{r}
var_data <- train_data$KitchenAbvGr %>% as.factor()
hist_category(var_data)
summary_category(var_data)
calc_corr(train_data$KitchenAbvGr)
```

I think kitchen quality is more important than this variable

### 51) KitchenQual: Kitchen quality

```{r}
var_data <- train_data$KitchenQual
hist_category(var_data)
summary_category(var_data)
```

Can be coverted to numeric

### 52) TotRmsAbvGrd: Total rooms above grade (does not include bathrooms)

For EDA, I am using treating this variable as a category variable

```{r}
var_data <- train_data$TotRmsAbvGrd %>% as.factor()
hist_category(var_data)
summary_category(var_data)
calc_corr(train_data$TotRmsAbvGrd)
```

### 53) Functional: Home functionality (Assume typical unless deductions are warranted)

```{r}
var_data <- train_data$Functional
hist_category(var_data)
summary_category(var_data)
```

Does not look very helpful. Can be coverted to numeric as with other quality values.

### 54) Fireplaces: Number of fireplaces

For EDA, I am using treating this variable as a category variable

```{r}
var_data <- train_data$Fireplaces %>% as.factor()
hist_category(var_data)
summary_category(var_data)
calc_corr(train_data$Fireplaces)
```

### 55) FireplaceQu: Fireplace quality

```{r}
var_data <- train_data$FireplaceQu
hist_category(var_data)
summary_category(var_data)
```

Can be converted to numeric and have to do a correlation check with the number of fireplaces variable

### 56) GarageType: Garage location

```{r}
var_data <- train_data$GarageType
hist_category(var_data)
summary_category(var_data)
```

Should try to convert this to numeric variable and see with the numbers as given in the description.

### 57) GarageYrBlt: Year garage was built
 
```{r}
plot_price_vs_var(train_data$GarageYrBlt)
summary_cont(train_data$GarageYrBlt)
```

Looks as important variable. Normalize it

### 58) GarageFinish: Interior finish of the garage

```{r}
var_data <- train_data$GarageFinish
hist_category(var_data)
summary_category(var_data)
```

Convert to numeric and check correlation with GarageYrBlt variable

### 59) GarageCars: Size of garage in car capacity

For EDA, I am using treating this variable as a category variable

```{r}
var_data <- train_data$GarageCars %>% as.factor()
hist_category(var_data)
summary_category(var_data)
calc_corr(train_data$GarageCars)
```

### 60) GarageArea: Size of garage in square feet

```{r}
plot_price_vs_var(train_data$GarageArea)
summary_cont(train_data$GarageArea)
```

### 61) GarageQual: Garage quality

```{r}
var_data <- train_data$GarageQual
hist_category(var_data)
summary_category(var_data)
```

Convert this to numeric and check correlation with other garage parameters.

### 62) PavedDrive: Paved driveway

```{r}
var_data <- train_data$PavedDrive
hist_category(var_data)
summary_category(var_data)
```

### 63)WoodDeckSF: Wood deck area in square feet

```{r}
plot_price_vs_var(train_data$WoodDeckSF)
summary_cont(train_data$WoodDeckSF)
```

### 64) OpenPorchSF: Open porch area in square feet

```{r}
plot_price_vs_var(train_data$OpenPorchSF)
summary_cont(train_data$OpenPorchSF)
```

### 65) EnclosedPorch: Enclosed porch area in square feet

```{r}
plot_price_vs_var(train_data$EnclosedPorch)
summary_cont(train_data$EnclosedPorch)
```

### 66) 3SsnPorch: Three season porch area in square feet

```{r}
plot_price_vs_var(train_data$X3SsnPorch)
summary_cont(train_data$X3SsnPorch)
```

### 67) ScreenPorch: Screen porch area in square feet

```{r}
plot_price_vs_var(train_data$ScreenPorch)
summary_cont(train_data$ScreenPorch)
```

### 68) PoolArea: Pool area in square feet

```{r}
plot_price_vs_var(train_data$PoolArea)
summary_cont(train_data$PoolArea)
```

Does not seem like a deciding variable

### 69) PoolQC: Pool quality

```{r}
var_data <- train_data$PoolQC
hist_category(var_data)
summary_category(var_data)
```

### 70) Fence: Fence quality

```{r}
var_data <- train_data$Fence
hist_category(var_data)
summary_category(var_data)
```

Same as other pool variable. Does not look like a deciding variable

### 71) MiscFeature: Miscellaneous feature not covered in other categories

```{r}
var_data <- train_data$MiscFeature
hist_category(var_data)
summary_category(var_data)
```

Does not look like important variable

### 72) MiscVal: $Value of miscellaneous feature

```{r}
var_data <- train_data$MiscVal
hist_category(var_data)
summary_category(var_data)
```

Not important variable

### 73) MoSold: Month Sold (MM)

```{r}
plot_price_vs_var(train_data$MoSold)
summary_cont(train_data$MoSold)
```

### 74) YrSold: Year Sold (YYYY)

```{r}
plot_price_vs_var(train_data$YrSold)
summary_cont(train_data$YrSold)
```

### 75) SaleType: Type of sale

```{r}
var_data <- train_data$SaleType
hist_category(var_data)
summary_category(var_data)
```

### 76) SaleCondition: Condition of sale

```{r}
var_data <- train_data$SaleCondition
hist_category(var_data)
summary_category(var_data)
```

### Sale Price

